A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2023-07-02 paper/fascd.tex
update Results section: Q1 results, better text

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.  Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.  Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.  For $u,v \in \cV$, $\eps>0$, and $\phi \in L^\infty(\Omega)$, define
\begin{equation}
\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} + b(u,v) - \ip{\phi}{v}. \label{eq:advectiondiffusion}
\end{equation}
\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
\begin{equation}
-\eps \grad^2 u + \bX\cdot \grad u = \phi.
\label{eq:pdeadvectiondiffusion}
\end{equation}
On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty + \|\phi\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.  By the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\dx + b(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
\end{example}

@@

Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, with $P_1$ elements and standard, global mesh refinement.  Here we show numerical results for the several VI Examples in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.  Our essential observations are that each V-cycle dramatically reduces the residual norm, and that a few F-cycles solves most Examples down to discretization error.

\begin{example} \label{ex:results:classical}
Consider the 2D classical, i.e.~unilateral and Laplacian, obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.  We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.   The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.  In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.  Note that $\psi\big|_{\partial \Omega} \le 0$ in both problems.

@@

Table \ref{tab:results:classical} shows the resulting iteration counts for the various finest meshes of $m_J$ nodes.  When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.  By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.  Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.  The spiral problem appears to be only slightly more challenging for these solvers.  The complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.

Each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.  For the F-cycle results, where three iterations suffice at all resolutions in either problem, this is close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  However, work of at most ten residual evaluations is often mentioned.} despite the low solution regularity.
\end{example}

%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets (black) for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:results:classical}
\end{figure}

% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py and edplay/results/drape.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}cccccc}
\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
 1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
 2 &    $9^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
 3 &   $17^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
 4 &   $33^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
 5 &   $65^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
 6 &  $129^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
 7 &  $257^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
 8 &  $513^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
 9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX
\end{tabular}
\bigskip
\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
\label{tab:results:classical}
\end{table}

\begin{example}  \label{ex:results:plap}
Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.  Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively.  Ellipticity is lost in the latter case.  For such $p$-Laplacian obstacle problems, well-posedness and ($p$-dependent) solution regularity are considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results, namely \pr{up} $=0$ and \pr{down} $=0$ in Algorithm \ref{alg:fascd}, and it emphasizes the need for robust smoothers.

Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.  The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.  For this experiment we used a coarsest mesh of $m_0=7$ nodes and an $m_9=3073$ finest mesh.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and again \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.

@@

Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the nonlinear FAS coarse correction VI subproblem solutions.  That is, each $y^j \in \mathcal{D}^j$ and $z^j \in \mathcal{U}^j$ is drawn, along with each level defect constraint $\underline{\phi}^j,\underline{\chi}^j$.

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
%   cd examples/
%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
\begin{figure}[ht]
\begin{center}
\input{tikz/imagesvcycle.tex}
\end{center}
\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ nodes.  Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.  The zero function (dotted) is the admissible initial iterate for $y^j$ and $z^j$.  Vertical scale varies.}
\label{fig:imagesvcycle}
\end{figure}

Iteration results are shown in Table \ref{tab:results:fastplap1d} for the fast diffusion $p=1.5$ case.  We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles (i.e.~Algorithm \ref{alg:fascd}).  F-cycles are even more efficient except when they fail entirely, but the failures arise from singular diffusivity.  If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').

Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.  V-cycles using the default settings, specifically three fixed RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.  Visualization shows an instability in which oscillations in the restricted residuals explode.  However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.  Even for the three-iteration solver, a single F-cycle suffices at every resolution.

@@

% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\ \cmidrule{1-8}
 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$
\end{tabular}
\bigskip
\caption{Iterations for V- and F-cycles on the $p=1.5$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$ and error.}
\label{tab:results:fastplap1d}
\end{table}

% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py and edplay/results/plap1d.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\ \cmidrule{1-6}
 1 &    $7$ &  1 &  1 &  1 & $8.7 \times 10^{-2}$ \\
 2 &   $13$ &  1 &  1 &  1 & $3.8 \times 10^{-2}$ \\
 3 &   $25$ &  2 &  1 &  1 & $1.6 \times 10^{-2}$ \\
 4 &   $49$ &  4 &  2 &  1 & $6.3 \times 10^{-3}$ \\
 5 &   $97$ & NC &  2 &  1 & $2.3 \times 10^{-3}$ \\
 6 &  $193$ & NC &  1 &  1 & $7.0 \times 10^{-4}$ \\
 7 &  $385$ & NC &  2 &  1 & $2.0 \times 10^{-4}$ \\
 8 &  $769$ & NC &  2 &  1 & $1.1 \times 10^{-4}$ \\
 9 & $1537$ & NC &  2 &  1 & $4.0 \times 10^{-5}$ \\
10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$
\end{tabular}
\bigskip
\caption{Iterations for V- and F-cycles on the $p=4.0$ problem in Example \ref{ex:results:plap}.  See text regarding V$_4$ and error.}
\label{tab:results:degenerateplap1d}
\end{table}

Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.  In fact, the above Examples, which were run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.  The next Example applies a parallel solver, and Example \ref{ex:results:sia} also demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.

\begin{example}  \label{ex:results:advdiff}
The advection-diffusion operator in Example \ref{ex:advectiondiffusion} is coercive when the velocity $\bX$ is divergence-free.  We now demonstrate FASCD F-cycles on such a VI problem over $\Omega=(-1,1)^d$ ($d=2,3$).  The 2D problem satisfies the coercivity hypotheses ($\bX = (7+5y,-5x)$ is divergence-free), but the 3D problem has a more general velocity ($\grad\cdot\bX \ne 0$); note $u=0$ on all of $\partial \Omega$ in both problems.  The discontinuous source term $\phi \in L^\infty(\Omega)$ has $\phi>0$ patches in half of the domain and $\phi<0$ values in the other half.  While the other Examples in this Section are unilateral, here we constrained the solution with upper and lower obstacles: $0 \le u \le 1$.  The $u=0$ and $u=1$ coincidence sets for the 2D problem are shown in Figure \ref{fig:results:advdiff}; the 3D sets are comparable in regularity.

% REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%   mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 2 -levs 7 -o poll2d.pvd
%   paraview poll2d.pvd
% to plot coincidence sets use ranges 0 <= u <= 0.00001 and 0.99999 <= u <= 1 and
% preset "xray" and invert colors (for u=0) or keep colors (u=1)
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
\end{center}
\caption{Coincidence sets (black) for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
\label{fig:results:advdiff}
\end{figure}

For the experiment we used a coarse mesh with $m_0=16^d$ nodes so as to minimally-resolve the coincidence set.  The smoother was two (fixed) RS Newton iterations,  with three preconditioned GMRES iterations to approximately solve the step equations.  The additive Schwarz method (ASM) preconditioner applied incomplete factorization (ILU) preconditioning on the blocks.  The coarse solver was RS Newton and LU, iterated to convergence.  Tolerances \id{rtol} $=10^{-5}$, \id{atol} $=10^{-9}$ were used in \eqref{eq:stoppingcriterion}.

>>>>

\begin{example}  \label{ex:advectiondiffusion}  Suppose $\partial\Omega$ is partitioned into Dirichlet and Neumann portions, i.e.~$\partial\Omega = \partial_D\Omega \cup \partial_N\Omega$, with $\partial_D\Omega$ of positive Hausdorff measure.  Let $\cV \subset W^{1,2}(\Omega)$ be the space of functions which are zero along $\partial_D\Omega$.  Consider a divergence-free velocity field $\bX$ on $\Omega$ satisfying the conditions of Lemma \ref{lem:advectionskew}, and assume that the flow is outward on the Neumann boundary: $\bX \cdot \bn \ge 0$ on $\partial_N\Omega$.  For $u,v \in \cV$, $\eps>0$, and $\phi \in L^\infty(\Omega)$, define
\begin{equation}
\ip{f(u)}{v} = \eps \left(\grad u, \grad v\right)_{L^2(\Omega)} + b(u,v) - \ip{\phi}{v}. \label{eq:advectiondiffusion}
\end{equation}
\pef{Why is $\phi \in L^\infty(\Omega)$? If that is the case, then should we have an $L^2$ inner product in the last term, rather than a dual pairing?}
With this operator, consider VI \eqref{eq:vi} with $\ell = 0$ for any closed and convex $\cK \subset \cV$. The interior condition, i.e.~PDE strong form, is the following linear advection-diffusion equation
\begin{equation}
-\eps \grad^2 u + \bX\cdot \grad u = \phi.
\label{eq:advectiondiffusionstrong}
\end{equation}
On the other hand, it is easy to see that $|\ip{f(u)}{v}| \le (\eps + \|\bX\|_\infty + \|\phi\|_\infty) \|u\| \|v\|$, thus that $f:\cK \to \cV'$ is continuous.  Lemma \ref{lem:advectionskew} says that the bilinear form $b(u,v)$ is skew-symmetric up to a nonnegative term.  By the outward flow assumption and the Poincar\'e inequality,
\begin{align*}
\ip{f(u)-f(v)}{u-v} &= \eps \int_\Omega |\grad u - \grad v|^2\dx + b(u-v,u-v) \\
                    &= \eps \int_\Omega |\grad u - \grad v|^2\dx + \frac{1}{2} \int_{\partial_N\Omega} (u-v)^2 \bX\cdot\bn \ge \eps C \|u-v\|^2.
\end{align*}
Thus $f$ is 2-coercive, and so the associated VI problem is well-posed.
\end{example}

@@

Algorithms \ref{alg:fascd} and \ref{alg:fascd-fcycle} were implemented in Python using the Firedrake library \cite{Rathgeberetal2016}, allowing $P_1$ or $Q_1$ elements and based on standard, global mesh refinement.  Here we show numerical results for the several example problems in Section \ref{sec:vi}, with a focus on the scaling of FASCD iterations as the mesh resolution increases.  Our essential observations are that each V-cycle dramatically reduces the residual norm, and that very few F-cycles solve most examples to within discretization error.

\begin{example} \label{ex:results:classical}
Consider the 2D classical, i.e.~unilateral and Laplacian, obstacle problem, namely Example \ref{ex:plaplacian} with $p=2$ and admissible set $\mathcal{K} = \{v \ge \psi\} \subset W^{1,2}(\Omega)$.  We consider two problems with very different coincidence (active) sets $\{x\in\Omega \,:\, u(x)=\psi(x)\}$, both over square $\Omega$.   The standard ``ball'' problem has an upper hemisphere obstacle $\psi$ \cite[Chapter 12]{Bueler2021}.  In the ``spiral'' problem from Example 7.1.1 in \cite{GraeserKornhuber2009}, $\psi$ is constructed to produce a more interesting coincidence set; see Figure \ref{fig:results:classical}.

@@

Table \ref{tab:results:classical} shows the resulting iteration counts for finest meshes of $m_J$ nodes.  When the RS Newton smoother is applied by itself as a solver (``RS only'') it shows rapidly-increasing iterations as expected, roughly proportional to resolution.  By contrast, V-cycle iterations increase slowly, perhaps logarithmically with the resolution.  Noting that for F-cycles we are counting the number of V-cycles after the ramp (Section \ref{sec:implementation}), we see that a constant number of cycles solves the problem.  Comparing the ball and spiral results, the complexity of the coincidence set or free boundary does not strongly-influence FASCD performance.  Note that each V-cycle does the work of about eight finest-level preconditioned Krylov (CG+IC) iterations.  Thus for the F-cycle results, where three iterations suffice at all resolutions, we see close to textbook multigrid efficiency\footnote{Textbook multigrid efficiency is imperfectly defined, but consider \cite{BrownSmithAhmadia2013}: ``an order of magnitude reduction in residual per iteration and solution of the fine-level system at a small multiple of the cost of a residual evaluation.''  Work of less than ten residual evaluations is often mentioned as the standard.} despite the low solution regularity.
\end{example}

%REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%$ timer python3 drape.py -prob ball|spiral -levs 9 -o ball|spiral.pvd -fascd_rtol 1.0e-6 -fascd_atol 1.0e-6 -fascd_cycle_type full
%$ paraview ball|spiral.pvd
%use preset "xray" and invert colors
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/ball-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/spiral-set.png}
\end{center}
\caption{Coincidence sets for the ball and spiral problems in Example \ref{ex:results:classical}.}
\label{fig:results:classical}
\end{figure}

% see git@bitbucket.org:pefarrell/fascd.git: examples/drape.py
% and https://github.com/bueler/mcd-extended: paper/results/drape.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}cccccc}
\multirow{2}{*}{\emph{levels}} & \multirow{2}{*}{$m_J$} & \multicolumn{3}{c}{\,\emph{ball}} & \multicolumn{3}{c}{\,\emph{spiral}} \\ \cmidrule(lr){3-5} \cmidrule(lr){6-8}
   &          & V-cycle & F-cycle & RS only & V-cycle & F-cycle & RS only \\ \cmidrule{1-8}
 1 &    $5^2$ &  1 &  1 &  1  &  1 &  1 &  1 \\
 2 &    $9^2$ &  2 &  1 &  2  &  2 &  1 &  2 \\
 3 &   $17^2$ &  2 &  2 &  3  &  3 &  2 &  4 \\
 4 &   $33^2$ &  3 &  2 &  6  &  3 &  2 &  7 \\
 5 &   $65^2$ &  3 &  2 &  8  &  4 &  3 &  9 \\
 6 &  $129^2$ &  4 &  2 & 17  &  4 &  3 & 16 \\
 7 &  $257^2$ &  4 &  2 & 33  &  4 &  3 & 27 \\
 8 &  $513^2$ &  5 &  2 & 65  &  5 &  3 & 48 \\
 9 & $1025^2$ &  5 &  2 & \XX &  5 &  3 & \XX \\
10 & $2049^2$ &  5 &  2 & \XX &  5 &  3 & \XX
\end{tabular}
\bigskip
\caption{Iterations for FASCD V-cycle and F-cycle algorithms in Example \ref{ex:results:classical}.  Smoother-only runs marked \XX\xspace were not attempted.}
\label{tab:results:classical}
\end{table}

\begin{example}  \label{ex:results:plap}
Recall Example \ref{ex:plaplacian}, the nonlinear $p$-Laplacian operator.  Here we consider a unilateral obstacle problem, in one dimension for visualizability, for both fast ($p=1.5$) and degenerate ($p=4.0$) diffusivity cases.\footnote{``Fast'' and ``degenerate'' refer to $D=|\grad u|^{p-2}\to \infty$ and $D\to 0$ as $|\grad u|\to 0$, respectively; ellipticity is lost in the latter case.  Well-posedness and ($p$-dependent) solution regularity for $p$-Laplacian obstacle problems is considered in \cite{ChoeLewis1991}.}  This example reveals the sometimes strong contrast between $V(1,0)$ and $V(0,1)$ results (\pr{up}, \pr{down} $=0$ in Algorithm \ref{alg:fascd}, respectively), and it emphasizes the need for robust smoothers.

Suppose $\Omega=(-3,3)$, $\psi(x) = -0.2|x|$, $\mathcal{K} = \{v \ge \psi\} \subset W^{1,p}(\Omega)$, and $\ell(v) = \int_\Omega g v$ where $g(x)=1$ for $|x|<1$ and $g(x)=-1$ for $|x|>1$.  The exact solution of VI \eqref{eq:vi} is easily computed for any $p$; see Figure \ref{fig:results:plap}.  For this experiment we used a coarsest mesh of $m_0=7$ nodes and an $m_9=3073$ finest mesh.  The default smoother uses three (fixed) iterations of RS Newton, with LU solution of the step equations.\footnote{This is an optimal $O(m_j)$ smoother in 1D.}  The coarse solver is the same but iterated to convergence.  The initial iterate was $w=\psi$, and \texttt{rtol} $= 10^{-6}$ and \texttt{atol} $= 10^{-12}$.  Figure \ref{fig:imagesvcycle} illustrates the major steps of an example V-cycle, showing each of the coarse-correction VI subproblems and solutions.

@@

% figures in the following visualization were generated using branch elbueler/simplefigs
% of git@bitbucket.org:pefarrell/fascd.git, and the following commands:
%   cd examples/
%   python3 plap1d.py -figures -fascd_max_it 2 -fascd_cycle_type multiplicative -mcoarse 2
\begin{figure}[ht]
\begin{center}
\input{tikz/imagesvcycle.tex}
\end{center}
\caption{Visualization of one FASCD V-cycle iteration for Example \ref{ex:results:plap}, with $p=1.5$, $J=3$, and a coarsest mesh of $m_0=3$ nodes.  Solutions (i.e.~$w^j$, $y^j$, $z^j$) are lines while obstacles ($\underline{\gamma}^j$, $\underline{\phi}^j$, $\underline{\chi}^j$) are strong dots.  The zero function is shown (dotted), but vertical scale varies.}
\label{fig:imagesvcycle}
\end{figure}

Iteration results are shown in Table \ref{tab:results:fastplap1d} for the fast diffusion $p=1.5$ case.  We see that $V(1,0)$ cycles, essentially Algorithm 4.7 in \cite{GraeserKornhuber2009} but with a stronger smoother, often do not converge.  By contrast, $V(0,1)$ cycles, using only up-smoothing in the larger sets $\mathcal{U}^j$ (Section \ref{sec:cdmultilevel}), are just as good as the default $V(1,1)$ cycles ($=$ Algorithm \ref{alg:fascd}).  F-cycles are even more efficient except when they fail entirely, but these failures arise from singular diffusivity.  If we regularize  $D=|\grad u|^{-0.5}$ to $D_\eps = \left(\eps + |\grad u|^2\right)^{-0.25}$, with $\eps=10^{-8}$, then a single F-cycle suffices at every resolution (column ``F$_\eps$'').

Iteration results for the (unregularized) degenerate diffusion $p=4.0$ case are in Table \ref{tab:results:degenerateplap1d}.  V-cycles using the default settings, specifically three fixed RS Newton iterations, do not converge on the finer meshes; the semi-smooth NCP residual norms \eqref{eq:stoppingcriterion} are unbounded in the failed runs.  Visualization shows an instability in which oscillations in the restricted residuals explode.  However, either a slight increase in smoother power, to four RS Newton iterations, or improved finest-level initial iterates through use of an F-cycle, stops the instability.  Even for the three-iteration smoother, a single F-cycle suffices at every resolution.

@@

% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 3mm}c@{\hskip 3mm}c@{\hskip 4mm}c@{\hskip 5mm}c@{\hskip 6mm}c}
\emph{levels} & $m_J$ & V(1,0) & V(0,1) & V(1,1) & F & F$_\eps$ & $\|$error$\|_\infty$ \\ \cmidrule{1-8}
 1 &    $7$ &  1 &  1 &  1 &  1 &  1 & $2.3\times 10^{-1}$ \\
 2 &   $13$ &  3 &  3 &  2 &  2 &  1 & $3.3 \times 10^{-2}$ \\
 3 &   $25$ & 12 &  4 &  2 &  1 &  1 & $9.1 \times 10^{-3}$ \\
 4 &   $49$ & NC &  4 &  2 &  1 &  1 & $3.2 \times 10^{-3}$ \\
 5 &   $97$ & NC &  3 &  3 &  1 &  1 & $5.5 \times 10^{-4}$ \\
 6 &  $193$ & NC &  3 &  3 & NC &  1 & $1.7 \times 10^{-4}$ \\
 7 &  $385$ & NC &  3 &  3 &  1 &  1 & $4.7 \times 10^{-5}$ \\
 8 &  $769$ & NC &  4 &  7 &  1 &  1 & $9.5 \times 10^{-6}$ \\
 9 & $1537$ & NC &  5 &  6 & NC &  1 & $3.6 \times 10^{-6}$ \\
10 & $3073$ & NC & 11 & 19 &  1 &  1 & $4.9 \times 10^{-7}$
\end{tabular}
\bigskip
\caption{Iterations for V- and F-cycles on the $p=1.5$ problem in Example \ref{ex:results:plap}.  NC $=$ did not converge after 50 iterations.  See text regarding F$_\eps$ and error.}
\label{tab:results:fastplap1d}
\end{table}

% see git@bitbucket.org:pefarrell/fascd.git: examples/plap1d.py
% and https://github.com/bueler/mcd-extended: paper/results/plap1d.sh|txt
\begin{table}[ht]
\begin{tabular}{lr@{\hskip 7mm}c@{\hskip 4mm}c@{\hskip 4mm}c@{\hskip 6mm}c}
\emph{levels} & $m_J$ & V & V$_4$ & F & $\|$error$\|_\infty$ \\ \cmidrule{1-6}
 1 &    $7$ &  1 &  1 &  1 & $8.7 \times 10^{-2}$ \\
 2 &   $13$ &  1 &  1 &  1 & $3.8 \times 10^{-2}$ \\
 3 &   $25$ &  2 &  1 &  1 & $1.6 \times 10^{-2}$ \\
 4 &   $49$ &  4 &  2 &  1 & $6.3 \times 10^{-3}$ \\
 5 &   $97$ & NC &  2 &  1 & $2.3 \times 10^{-3}$ \\
 6 &  $193$ & NC &  1 &  1 & $7.0 \times 10^{-4}$ \\
 7 &  $385$ & NC &  2 &  1 & $2.0 \times 10^{-4}$ \\
 8 &  $769$ & NC &  2 &  1 & $1.1 \times 10^{-4}$ \\
 9 & $1537$ & NC &  2 &  1 & $4.0 \times 10^{-5}$ \\
10 & $3073$ & NC &  2 &  1 & $1.5 \times 10^{-5}$
\end{tabular}
\bigskip
\caption{Iterations for V- and F-cycles on the $p=4.0$ problem in Example \ref{ex:results:plap}.  See text regarding V$_4$ and error.}
\label{tab:results:degenerateplap1d}
\end{table}

Solvers implemented in Firedrake can exploit parallel solvers from the PETSc library \cite{Balayetal2023}.  The above Examples, run in serial so as to fix the smoother,\footnote{The action of parallel preconditioners generally depends on the number of processes \cite[for example]{Bueler2021}.} also run efficiently in parallel.  The next Example applies a parallel solver, and Example \ref{ex:results:sia} demonstrates the parallel weak scaling \cite{Bueler2021} of FASCD.

\begin{example}  \label{ex:results:advdiff}
We applied FASCD F-cycles to the advection-diffusion VI problem in Example \ref{ex:advectiondiffusion}.  The domain was $\Omega=(-1,1)^d$, for $d=2,3$, diffusivity was $\eps=0.1$ (equation \eqref{eq:advectiondiffusionstrong}), and there were zero Dirichlet boundary conditions.  The 2D problem satisfied the coercivity hypothese with a divergence-free velocity field $\bX = (7+5y,-5x)$, but in 3D we set $\bX = (7+5y,-5x,2z)$ so $\grad\cdot\bX \ne 0$.  The discontinuous source term $\phi \in L^\infty(\Omega)$ satisfied $\phi\ge 0$ in half of the domain, with positive circular patches, while values were negative in the other half.  While the other Examples in this Section are unilateral, here the solution was box-constrained: $0 \le u \le 1$.  Figure \ref{fig:results:advdiff} shows coincidence sets for the 2D problem; the 3D sets are comparable in regularity and general pattern.

% REGENERATE using git@bitbucket.org:pefarrell/fascd.git, in examples/
%   mpiexec -n 16 --map-by core --bind-to hwthread python3 pollutant.py -dim 2 -levs 7 -o poll2d.pvd
%   paraview poll2d.pvd
% to plot coincidence sets use ranges 0 <= u <= 0.00001 and 0.99999 <= u <= 1 and
% preset "xray" and invert colors (for u=0) or keep colors (u=1)
\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.35\textwidth]{fixfigs/poll2d-zero-set.png} \qquad \includegraphics[width=0.35\textwidth]{fixfigs/poll2d-one-set.png}
\end{center}
\caption{Coincidence sets for the 2D bi-obstacle problem in Example \ref{ex:results:advdiff}: $u=0$ (left) and $u=1$ (right).}
\label{fig:results:advdiff}
\end{figure}
